---
title: "ECON 339 - Tuesday Week 2"
date: "April 9 , 2024"
author: "Matina Lampsas"
format: 
   html:
     embed-resources: true
     code-tools: true
     toc: true
     code-fold: show
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

```{r setup}
library(tidyverse)
library(readxl)
options(scipen = 999)
```

# Read in the data
```{r}
#read in the data for the college 
college_data <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "College", 
                     range = "A1:G117",
                      n_max = 118
                      )
```

You can also use the import data feature. But to keep in tidy format use the read_xlsx function

# Find the mean in Base R:
```{r}
# find the mean of the cost
mean(college_data$Cost)
#find the mean of the grad
mean(college_data$Grad)

```


# Find the mean but using DPLYR - a tidy format of the above: 
```{r}
# Better way to do the above:
# Mean of Cost and grad
college_data %>%
  summarise(mean_cost = mean(Cost, na.rm = T),
            mean_grad = mean(Grad, na.rm = T))
```

# Make the city variable 
Assign a 1 to schools that are a city, and a 0 to cities that are non city. 
```{r}
#make the city variable: 
college_data <- college_data %>% 
  mutate(City = if_else(Location == "City", 1, 0))

college_data
```

# Run a Regression: 
we are goinng to run the same regression as we did in excel: 
```{r}
library(broom)
# run a regression model 
lm_model <- lm(Earnings ~ Cost+Grad+Debt+City, data = college_data)
summary(lm_model)
coefs <- broom::tidy(lm_model)
coefs

```
The function for the linear model is : Function is y= (`r as.numeric(coefs[1,2])`) +(`r as.numeric(coefs[2,2])`)x1 + (`r as.numeric(coefs[3,2])`)x2+ (`r as.numeric(coefs[4,2])`)x3 + (`r as.numeric(coefs[5,2])`)x4

# Use Predict to estimate what the value of earnings
Given the Cost of a school to be $25,000, the Graduation rate of 60%, Debt of 80%, and the school being in a city. Calculate the estimated earnings of the student after graduation. Use the lm created to estimate this. 
```{r}
# predict 
predict(lm_model, data.frame(Cost = 25000, Grad = 60, Debt = 80, City = 1))
```


# Fitted vs Residuals plot: 

```{r}
lm_model %>%
  ggplot(aes(x = .fitted, y = .resid )) +
  geom_point()+
  geom_smooth(method = "lm")
```

