---
title: "ECON 339 - Tuesday/Thursday Week 8"
date: "May 21, 2024"
author: "Matina Lampsas"
format: 
   html:
     embed-resources: true
     code-tools: true
     toc: true
     code-fold: show
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
  
---

# Setup
```{r setup}
library(tidyverse)
library(readxl)
library(car)
options(scipen = 999)

```
# Tuesday Notes
## Read in the Data 
```{r}
#read in the data for the wages 
satisfaction_data <- read_xlsx(here::here("data",
                     "Data for Weeks 7-10.xlsx"), 
                      sheet = "Satisfaction"
                     )
```

## Practice Problem: 
### Estimate the true and used models: 
```{r}
true_model <- lm(Satisfaction ~ Salary+Hours, data = satisfaction_data)
used_model <- lm(Satisfaction ~ Salary, data = satisfaction_data)

summary(true_model)
summary(used_model)

```
Looking at the Summaries of the True and Used models we can better understand the bias and if Endogeneity is present in the model. 

True Model: 
There is a positive correlation between salary and satisfaction, but there is a negative correlation between hours and satisfaction. This shows a negative bias


## Practice Problem - Heteroskedasticity
 Consider a simple regression model that relates monthly sales (Sales in $1,000s) from a 
chain of convenience stores with the square footage (Sqft) of the store. (Data: Stores)
 • You would expect sales to vary more as square footage increases. For instance, a 
small convenience store is likely to include only bare essentials for which there is a 
fairly stable demand. A larger store, on the other hand, may include specialty items, 
resulting in more fluctuation in sales
```{r}
#read in the data for the wages 
stores_data <- read_xlsx(here::here("data",
                     "Data for Weeks 7-10.xlsx"), 
                      sheet = "Stores"
                     )
```
### Make a residuals vs fitted plot: 
In order to better understand if we have heteroskedasticity present, lets make a residuals vs fitted plot: 
```{r}
stores_lm <- lm(Sales ~ Sqft, data = stores_data)

# Add residuals and the square root of the predictor variable to the data frame
stores_data <- stores_data %>%
  mutate(residuals = resid(stores_lm))

stores_data %>% 
  ggplot(aes(x = Sqft, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +  # Optional: add a smooth line
  labs(title = "Residuals vs. Square Footage",
       x = "Squarefootage",
       y = "Residuals") +
  theme_bw()+
  theme(axis.text.x = element_text(face = "bold"))
```
```{r}
# load in the libraries: 
library(zoo)
library(sandwich)
```
### To calculate robust standard error: 
```{r}
coeftest(stores_lm, vcov. = vcovHC(stores_lm, type = "HC1"))
```

## Practice Problem 3: 
```{r}
# load in the data 
#read in the data for the college 
college_data <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "College", 
                     range = "A1:G117",
                      n_max = 118
                      )
```

### Make your linear regression model: 
```{r}
college_lm <- lm(Earnings ~ Cost + Grad + Debt + City, data = college_data)
summary(college_lm)
```
### Make your graphs: 
```{r}

# Add residuals and the square root of the predictor variable to the data frame
college_data <- college_data %>%
  mutate(residuals = resid(college_lm))

long_data <- college_data %>%
  pivot_longer(cols = c(Cost, Grad, Debt ,City), names_to = "variable", values_to = "value")


long_data %>% 
  ggplot(aes(x = value, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +  # Optional: add a smooth line
  labs(title = "Residuals vs. Square Footage",
       x = "Squarefootage",
       y = "Residuals") +
  theme_bw()+
  facet_wrap(~ variable, scales = "free_x") + # Create a separate panel for each predictor variable
  theme(axis.text.x = element_text(face = "bold"))
```

