---
title: "ECON 339 - Tuesday Week 5"
date: "April 30 , 2024"
author: "Matina Lampsas"
format: 
   html:
     embed-resources: true
     code-tools: true
     toc: true
     code-fold: show
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

# Setup
```{r setup}
library(tidyverse)
library(readxl)
options(scipen = 999)

```

# Read in the data
```{r}
#read in the data for the college 
MIS_data <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "Salary_MIS"
                     )
```

#Run linear regression model with and without the interaction variable
```{r}
# run a lm without the interaction varaible
lm_1 <- lm(Salary~ GPA + MIS + Statistics, data = MIS_data)
# run a lm with the interaction variable -- the interaction variable is MIS*Statistics
lm_interact <- lm(Salary~ GPA + MIS + Statistics + MIS*Statistics, data = MIS_data)
summary(lm_1) ; summary(lm_interact)
```

```{r}
# Example of predicting a variable using the interaction regression w/ two dummy variables
example <- predict(lm_interact, data.frame(GPA = 3, MIS = 1, Statistics = 1))
example
```
# Interaction of Numerical and Dummy variables: 
Here we have a dataset called Gender_Gap. It includes: 
- Salary: the salary (in dollars) of the individual
- Size: the size of the firm the project manager works at 
- Experience: the years of experience the person has
- Female: a dummy variable that assigns a value of 1 if the individual is a female and a 0 if they are not a female
- Grad: assigns 1 to those who went to grad school, and 0 to those who did not
## Import the Data
```{r}
#read in the data for the college 
gender_gap <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "Gender_Gap"
                     )
```

## Now run a linear regression w/ interaction variables: 
```{r}
gender_gap_lm <- lm(Salary ~ Size + Experience + Female + Grad + Female*Experience + Female*Grad, data = gender_gap)

```
Now that we read in the data and ran a linear model, we can see how experience interacts with gender to affect salary. Also we can see how gender interacts with graduate degree to influence salary

## Visualize the results: 
```{r}
msize = as.numeric(mean(gender_gap$Size))
exp <- 1:20
female <- map(.x = 1:20, 
              .f = ~predict(gender_gap_lm, data.frame(Size = msize, Experience = .x, Female = 1, Grad = 1)))
male <- map(.x = 1:20, 
              .f = ~predict(gender_gap_lm, data.frame(Size = msize, Experience = .x, Female = 0, Grad = 1)))

```

# Practice Problem
## Import the Data
```{r}
#read in the data for the college 
conversion <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "Conversion"
                     )
```

## Part A: 
Create a dummy variable labeled Training, that assigns a value of 1 for employees who 
participate in training and 0 for those who do not. Estimate and interpret the effect of 
Experience and Training on the lead conversion rate.
```{r}
# creating our dummy variable
conversion <- conversion %>% 
  mutate(Training = if_else(Participation == "Yes", 1, 0))
  
  # run the model with rate as the response and experience/training as the explanatory 
 conversion_lm_1 <- lm(Rate ~ Experience + Training , data = conversion)
 broom::tidy(conversion_lm_1)

```
Now that we read in the data and ran a linear model, we can see how experience interacts with gender to affect salary. Also we can see how gender interacts with graduate degree to influence salary

## Part B:
It is believed that training participation may be especially effective for managers with fewer 
years of experience. Extend the above model by incorporating the appropriate interaction 
variable.
```{r}
# Run the lm with the interaction variable included 
 conversion_lm_2 <- lm(Rate ~ Experience + Training + Experience*Training, data = conversion)
 broom::tidy(conversion_lm_2)
```


## Part C: 
Find the preferred model and use the estimated model to predict the lead conversion rate 
for an employee with 5 years of experience who has participated in training. What is the 
corresponding rate with 20 years of experience

### 5 & 20 year predictions: 
```{r}
five_yr_predict <- predict(conversion_lm_2, data.frame(Experience = 5, Training = 1))
twnty_yr_predict <- predict(conversion_lm_2, data.frame(Experience = 20, Training = 1))
five_yr_predict ; twnty_yr_predict
```
The prediction for an employee with 5 years of experience who has participated in the training is: `r (five_yr_predict)`

The prediction for an employee with 20 years of experience who has participated in the training is: `r (twnty_yr_predict)`

### Predict values: 
Predict the values of Employees ranging from 1-20 years experience who have and have not completed the training. 
After predicting the values, plot them to show a line representing the two scenarios. 
```{r}
# prediction values for those who are trained
train_predict <- map_df(.x = 1:20,
       .f = ~predict(conversion_lm_2, data.frame(Experience = .x, Training = 1)))
# prediction values for those who aren't trained
not_train_predict <- map_df(.x = 1:20,
       .f = ~predict(conversion_lm_2, data.frame(Experience = .x, Training = 0)))
# data frame with all of our data
conversion_predict <- data.frame(Experience = 1:20, Predictions_1 = train_predict, Prediction_2 = not_train_predict)

# Visualization - plot both lines on the same graph to visualize the predicted data
conversion_predict %>% 
  rename(prediction_trained = X1,
         prediction_not_trained =X1.1 ) %>% 
  ggplot(aes(x = Experience), color = prediction_trained)+
  geom_smooth(aes(y = prediction_trained, color = "Trailed"), method = "lm")+
  geom_smooth(aes(y = prediction_not_trained, color = "Not Trained"), method = "lm")+
  labs(x = "Years of Experience", y = "Salary", colour = "Key", title = "Linear Regression of Prediction of Salary Based on Years of Experience")

```



