---
title: "ECON 339 - Tuesday/Thursday Week 7"
date: "May 14, 2024"
author: "Matina Lampsas"
format: 
   html:
     embed-resources: true
     code-tools: true
     toc: true
     code-fold: show
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
  
---

# Setup
```{r setup}
library(tidyverse)
library(readxl)
library(car)
options(scipen = 999)

```
# Tuesday Notes
## Read in the Data 
```{r}
#read in the data for the wages 
multi_data <- read_xlsx(here::here("data",
                     "Data for Weeks 7-10.xlsx"), 
                      sheet = "Multi"
                     )
```

 Example: A regression model was developed to analyze the relationship between employee performance scores 
(ranging from 1 to 100) and several key factors, including whether the employee has a college degree (Yes/No), 
their age (in years), work experience (in years), and the percentage of sales targets achieved (Targets, in %). 
Explore the issue of multicollinearity in this application. If it is found to be severe, develop the appropriate 
model to capture the influence of the predictor variables correctly. (Data: Multi)

```{r}
multi_data <- multi_data %>% 
  mutate(d1 = if_else(College == "Yes", 1,0))

# replicate the lm from the slides

multi_lm1 <- lm(Score ~ Age + Experience + Targets + d1, data = multi_data)
multi_lm2 <- lm(Score ~ Experience + Targets + d1, data = multi_data)
multi_lm3 <- lm(Score ~ Age + Targets + d1, data = multi_data)

# summary(multi_lm1)
# summary(multi_lm2)
# summary(multi_lm3)

# extract the r-squared from the model
Rsqa1 <- summary(multi_lm1)$r.squared
Rsqa2 <- summary(multi_lm2)$r.squared
Rsqa3 <- summary(multi_lm3)$r.squared


# calculating the varaince inflation factor for the model
vif(multi_lm1)
vif(multi_lm2)
vif(multi_lm3)

# another way to calculate the correlation
# cor(multi_data[,3:5])
# multi_data %>%
#   select(Age, Experience, Targets) %>%
#   cor()

```
## Practice Problem 
Ara Lily wishes to analyze the gender gap in salaries of project 
managers. She has access to the salary (in $1,000s) for 200 project managers in small- to 
middle-sized firms in the Boston area. In addition, she has data on the firm size, the 
manager’s age and experience (in years), whether the manager is a female (Female 
equals 1 if female, 0 otherwise). Use all observations to answer the following questions. 
[Data: Manager]
### Read in the data
```{r}
manager_data <- read_xlsx(here::here("data",
                     "Data for Weeks 7-10.xlsx"), 
                      sheet = "Manager"
                     )
```

Estimate and report a model where the dependent variable is Salary and the explanatory 
variables include Size, Age, Experience, and Female.  Discuss the possible problem of 
multicollinearity in the data
```{r}
# replicate the lm from the slides
# if you do two models: 1 without 

# If you do two models: 
#   Model 1: with age and expirence
#   Model 2: only expirence, not age
# You will find the model 2 will have a higher R^2 therefore it is better to omit age when calculating 

Model1 <- lm(Salary ~ Size + Age + Exper + Female, data = manager_data)
summary(Model1)
cor(manager_data[,2:5])
#install.packages("car")

Model2 <- lm(Salary ~ Size + Exper + Female, data = manager_data)
summary(Model2)
Model3 <- lm(Salary ~ Size + Age + Female, data = manager_data)
summary(Model3)


library(car)

Rsqa1 <- summary(Model1)$r.squared
vif(Model1)

manager_data %>%
  select(Size, Age, Exper, Female) %>%
  cor()

```

Model 2 is better because it has higher R^2