---
title: "ECON 339 - Tuesday Week 3"
date: "April 16 , 2024"
author: "Matina Lampsas"
format: 
   html:
     embed-resources: true
     code-tools: true
     toc: true
     code-fold: show
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

# Setup
```{r setup}
library(tidyverse)
library(readxl)
options(scipen = 999)

```

# Read in the data
```{r}
#read in the data for the college 
retail_data <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "Retail", 
                     range = "A1:D41",
                      n_max = 44
                      )
```

# Create the Dummy Variables: 
```{r}
#make 3 dummy variables: 
retail_data <- retail_data %>% 
  mutate(d1 = if_else(Quarter == 1, 1,0),
         d2 = if_else(Quarter == 2, 1,0),
         d3 = if_else(Quarter == 3, 1,0))

```

# Make the regression Model
```{r}
retail_lm <- lm(Sales~GNP+d1+d2+d3, data = retail_data)
predict(retail_lm, data.frame(GNP = 18000, d1 = 0, d2= 1, d3 = 0))
coefs <- broom::tidy(retail_lm)
summary(retail_lm)
```
## Interpretation of Slope of d4 :
for every 1 unit increase of GNP, the sales will increase by (`r as.numeric(coefs[2,2])`) billion in Q4


# Predict 
## What are the predicted sales in quarter 2 if GNP is $18,000 (in billions)?
```{r}
predict(retail_lm, data.frame(GNP = 18000, d1 = 0, d2= 1, d3 = 0))
```

## Does using a different quarter for reference impact the inference?
Yes, because the reason d1,d2,d3 are negative is because we use d4 as a reference and the sales are higher in d4 than in other quarters. If we change the reference variable than yes it will impact the inference since the slopes of the dummy variables will change. 

# Homework - probelm: 
First subset the House_Price data by the college town of Ames (Iowa State 
University) or the college town of Lincoln (University of Nebraska). Also, only include Single 
Family houses. You should obtain 612 observations between these two towns, of which 209 are 
in Ames and 403 are in Lincoln.
## Read in the housing data
```{r}
house_price_data <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "House_Price"
                      )
```

## Question 1:
 Find the averages for the sale amount, number of bedrooms, number of bathrooms, square 
footage, and lot size in both campus towns.
```{r}
avg_ames <- house_price_data %>% 
  filter(Town == "Ames, IA") %>% 
  summarise(across(.cols = Sale_amount:Sqft_lot,
                   .fns = ~mean(.x)))
avg_lincoln <- house_price_data %>% 
  filter(Town == "Lincoln, NE") %>% 
  summarise(across(.cols = Sale_amount:Sqft_lot,
                   .fns = ~mean(.x)))
```
## Question 2/3: 
2. Make the two linear models from the slides
3. Which of the above models is better for making predictions?

```{r}
house_price_data <- house_price_data %>% 
  filter(Town == "Lincoln, NE" |
        Town == "Ames, IA") %>% 
  mutate(ames = if_else(Town == "Ames, IA", 1, 0),
         newer = if_else(Build_year> 2000, 1,0 ))

lm_1 <- lm(Sale_amount~Beds+Baths+Sqft_home+Sqft_lot+ames, data = house_price_data)
lm_2 <- lm(Sale_amount~Beds+Baths+Sqft_home+Sqft_lot+ames+newer, data = house_price_data)

broom::tidy(summary(lm_1))
broom::tidy(summary(lm_2))
summary(lm_1)
summary(lm_2)
```

Model 2 is better since it has a higher adjusted R^2. 
