---
title: "ECON 339 - Tuesday Week 4"
date: "April 23 , 2024"
author: "Matina Lampsas"
format: 
   html:
     embed-resources: true
     code-tools: true
     toc: true
     code-fold: show
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

# Setup
```{r setup}
library(tidyverse)
library(readxl)
options(scipen = 999)

```

# Read in the data
```{r}
#read in the data for the college 
JJ_data <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "J&J"
                     )
```

# Individual Test For Significance: 
## First create a regression model: 
### Q1 - a. At the 5% significance level, is the beta coefficient less than one? 

```{r}
jj_lm <- lm(`R-Rf`~`Rm-Rf`, data = JJ_data)
# summary(jj_lm)
coefs <- broom::tidy(jj_lm)
coefs
# pt(-1.796,58,lower.tail = TRUE)
# pt(1.796,58,lower.tail=FALSE)
df <- as.numeric(df.residual(jj_lm))
t_58 <- as.numeric((coefs[2,2]-1)/coefs[2,3])

pt(t_58, df, lower.tail = F)

```
Since the P-value is 0.038, it is less than .05 so we reject the null hypothesis that johnson and johnson is less risky than the market

### Q2- At the 5% significance level, are there abnormal returns? In other words, is 
the alpha coefficient significantly different from zero?
No there are no Abnormal Returns since the p-value for the intercept is too high 

# Practice Problem 
 Reconsider the Retail data to estimate a linear regression model with Sales as the dependent 
variable and GNP along with the relevant dummy variables, to capture seasonal variations, as 
explanatory variables. At the 5% significance level, test if the sales are higher in the 2nd quarter 
than the 3rd quarter. 
```{r}
retail_data <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "Retail", 
                     range = "A1:G41"
                     )
```

```{r}
# make the linear regression 
retail_lm <- lm(Sales~GNP+d1+d2+d4, data = retail_data)
summary(retail_lm)
```
```{r}
b <- retail_lm$coefficients[4]
se <- sqrt(diag(vcov(retail_lm)))[4]
n <- nrow(retail_data); k=3
t <- (b)/se; t
pt(t, (35), lower.tail = T)
```

