---
title: "ECON 339 - Tuesday Week 6"
date: "May 7, 2024"
author: "Matina Lampsas"
format: 
   html:
     embed-resources: true
     code-tools: true
     toc: true
     code-fold: show
editor: source
execute: 
  error: true
  echo: true
  message: false
  warning: false
---

# Setup
```{r setup}
library(tidyverse)
library(readxl)
options(scipen = 999)

```

# Read in the Data 
```{r}
#read in the data for the wages 
wages_data <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "Wages"
                     )
```

# In Class Example:
## Make a scatter plot: 
Plot Wage against Age and evaluate whether a linear or quadratic regression 
model better captures the relationship. Verify your choice by using the 
appropriate goodness-of-fit measure.
```{r}
wages_data %>% 
  ggplot(aes(x = Age, y = Wage))+
  geom_jitter()+ 
  geom_smooth()+ 
  geom_smooth(method = "lm", color = "red")
```
* On the exam, he will ask us to do graphs

Use the appropriate model to predict hourly wages for someone with 16 
years of education and age equal to 30, 50, or 70
## Conduct/predict using lm: 
```{r}
wages_data <- wages_data %>% 
  mutate(age_sqrd = Age^2)
wages_lm <- lm(Wage ~ Age + Educ ,data = wages_data)
summary(wages_lm)
wages_qm <- lm(Wage ~ Educ+ Age + I(Age^2), data = wages_data)
summary(wages_qm)
```
Predict Vals: 
```{r}
predict(wages_qm, data.frame(Educ=16, Age=30))
predict(wages_qm, data.frame(Educ=16, Age=50))
predict(wages_qm, data.frame(Educ=16, Age=70))

```
NOTE TO SELF --> FIND A WAY TO DO THIS WITHOUT THE I() THING
According to the model, at what age will someone with 16 years of 
education attain the highest wages?

```{r}
coeffs <- coefficients(wages_qm)
predict <- map2_df(.x = 20:80,
                   .y= 16,
                   .f = ~predict(wages_qm, data.frame(Educ = .y, Age = .x)))
predict %>%
  mutate(age = 20:80) %>%
  rename(pred_wage = `1`) %>%
  ggplot(aes(x = age, y = pred_wage)) +
  geom_jitter() +
  geom_smooth(color = "blue", aes(group = "Quadratic Model")) +  # Labeling the blue line as quadratic
  geom_smooth(method = "lm", color = "red", aes(group = "Linear Model")) +  # Labeling the red line as linear
  labs(title = "Predicted Quad/Linear graph", x = "Age", y = "Predicted Wage", color = "Model Key")
```

# Practice Question: 
## Import the Data: 
```{r}
#read in the data for the tech sales 
techsales_data <- read_xlsx(here::here("data",
                     "Data for Weeks 2-7.xlsx"), 
                      sheet = "TechSales"
                     )
```

## Subsest the data and make dummy variables: 
First subset the TechSales data to include only hardware product. You should obtain 9860 
observations. Use College and Personality variables to construct appropriate dummy 
variables. 
```{r}
filtered_tchsales <- techsales_data %>% 
  filter(Business == "Hardware") %>% 
  mutate(d1 = if_else(Personality == "Diplomat",1,0),
         d2 = if_else(Personality == "Explorer", 1,0),
         d3 = if_else(Personality == "Analyst",1,0),
         d4 = if_else(Personality == "Sentinel",1,0),
         d5 = if_else(College =="Yes", 1,0))

```
Compare the linear model (Model 1) that uses Age, Female, College, Analyst, 
Diplomat, Explorer, and Certificates as predictor variables. Model 2 extends Model 1 by 
including age-sq as an additional predictor variable
```{r}
# make a model that isn't quadratic
model_1 <- lm(NPS ~ Age + Female +Certificates + d1 + d2 + d3 + d5, data = filtered_tchsales)
# make a quadratic model that considers Age ^2
model_2 <- lm(NPS ~ Age + Female +Certificates+ d1 + d2 + d3 + d5 + I(Age^2), data = filtered_tchsales)
# summarize and compare the models
summary(model_1) ; summary(model_2)
```

• Use model selection criteria to find the appropriate model for the analysis. 
• Interpret your results with reference to well-formatted figures and tables.

The second model will be chosen since the R^2 value is larger and theremore more accurate

## Second Part of Practice Question:
Create a plot to highlight the quadratic effect of age. For the plot, use age between 30-60, using a female graduate analyst with 3 certificates

```{r}
# predict the values 
predict_2 <- map_df(.x = 30:60,
                    .f = ~predict(model_2, data.frame(Age = .x, 
                                                      Female = 1, 
                                                      d1 = 0,
                                                      d2 = 0,
                                                      d3 = 1,
                                                      d5 = 1,
                                                      Certificates = 3
                                                      )
                                  )
                    )

predict_2 %>%
  # add an age column to the data
  mutate(age = 30:60) %>%
  #rename the column from the predicted varaible
  rename(pred_NPS = `1`) %>%
  # make the visualization
  ggplot(aes(x = age, y = pred_NPS)) +
  geom_jitter() +
  geom_smooth(color = "blue",aes( fill = "Quadratic Model")) +  # Labeling the blue line as quadratic
  geom_smooth(method = "lm", color = "red", aes(fill = "Linear Model")) +  # Labeling the red line as linear
  labs(title = "Predicted Quad/Linear graph", x = "Age", y = "Predicted Wage", color = "Model Key") 

```
 
